Couple selection in UKBB

### 2021-11-23 analysis/couple_selection.Rmd
add phenotype and couple selection schematic links
wflow_publish("analysis/couple_selection.Rmd")

# UKBB full sample

@@

# European, genetic, consenting sample

@@

- `excess.relatives=0`
- `putative.sex.chromosome.aneuploidy=0`
- `in.white.british.ancestry.subset==1`

@@

- Raw phenotypes in the SGG database were processed according the the `PHESANT` pipeline. 
- For details on the processing by PHESANT see: https://github.com/astheeggeggs/PHESANT. 
- The pipeline is not the simplest to follow and does a lot of referring to other variables and has been modified slightly based on the data we have. 
- The main thing to note is all continuous variables were transformed to a normal distribution using an inverse normal rank transformation.
- Each phenotype file contains ID, age, and sex columns (the first 3).
- Note that this filtering procedure does not necessarily remove related individuals. For this, please use the function 'ukb_gen_samples_to_remove()' from the [UKBB r-package](https://cran.r-project.org/web/packages/ukbtools/vignettes/explore-ukb-data.html).

# Houshold pairs

@@

# Unrelated pairs

@@

# Heterosexual pairs

@@

# Couple pairs

>>>>

An overview of couple selection and exclusions can be seen [here](assets/couple_selection.png).

# UKBB full sample.

@@

# European, genetic, consenting sample.

@@

- `excess.relatives=0`
- `putative.sex.chromosome.aneuploidy=0`
- `in.white.british.ancestry.subset==1`
- `used.in.pca.calculation==1`

@@

- Raw phenotypes in the SGG database were processed according the the `PHESANT` pipeline. 
- For details on the processing by PHESANT see: https://github.com/astheeggeggs/PHESANT. 
- The pipeline is not the simplest to follow and does a lot of referring to other variables and has been modified slightly based on the data we have. 
- The main thing to note is all continuous variables were transformed to a normal distribution using an inverse normal rank transformation.
- Each phenotype file contains ID, age, and sex columns (the first 3).
- Note that this filtering procedure does not necessarily remove related individuals. For this, please use the function `ukb_gen_samples_to_remove()` from the [UKBB r-package](https://cran.r-project.org/web/packages/ukbtools/vignettes/explore-ukb-data.html).

# Houshold pairs.

@@

# Unrelated pairs.

@@

# Heterosexual pairs.

@@

# Couple pairs.
